Improving \textsl{Gaia} parallax precision \\ with a data-driven model of stars

### 2017-06-06 paper/paper.tex
fixed quotes

This change in precision is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes in the variance. The left panel shows the natural log of the fractional change in variance as a function of $(J-K_s)^C$ color, with the $1\sigma$ and $2\sigma$ contours over plotted.
The points are colored by the natural log of their fractional change in variance (the y-axis of the left panel) in both panels to help guide your eye for the right panel. The regions of \cmd\ space with the greatest improvements are shown in black, and the regions with the lowest improvements are shown in yellow.
The giants show obvious large improvements, as well as some main sequence stars.
Regions of large improvement are those with the largest photometric and parallax errors, which are the faintest observed objects.
This is a fairly standard result which we illustrated in the introductory section of this paper: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
Stars with a color $(J-K_s)^C \sim 0.6$ show a decrease in precision due to the width of the prior there; the more vertical structure in the red giant branch of the \cmd.

@@

We select sources from our catalog within a $1~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \gaia\ within this window on the sky.
Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not at all obvious.
The posterior PDFs, shown in the bottom row, are more sharply peaked showing an increase in precision.
The posterior PDFs also have modes that are more similar to the previously determined distance to the cluster (vertical lines) for far more stars.
This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{distancesM67.png}
\caption{Proof of concept: parallax inference to the open cluster M67.
The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $1~{\rm deg}^2$ window centered on the position of M67.
Our prior, visualized as a mixture of Gaussians, is shown in green.
The upper right panel shows the \gaia\ parallax likelihoods in black, with the "true" distance to M67, determined by previous methods, bracketed by the red lines.
Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
The posterior PDFs are much more clustered and sharply peaked around the "true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
\label{fig:m67}
\end{figure}

>>>>

This change in precision is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes in the variance. The left panel shows the natural log of the fractional change in variance as a function of $(J-K_s)^C$ color, with the $1\sigma$ and $2\sigma$ contours over plotted.
The points are colored by the natural log of their fractional change in variance (the y-axis of the left panel) in both panels to help guide your eye for the right panel. The regions of \cmd\ space with the greatest improvements are shown in black, and the regions with the lowest improvements are shown in yellow.
The giants show obvious large improvements, as well as some main sequence stars.
Regions of large improvement are those with the largest photometric and parallax errors, which are the faintest observed objects.
This is a fairly standard result which we illustrated in the toy model above: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
Stars with a color $(J-K_s)^C \sim 0.6$ show a decrease in precision due to the width of the prior there; the more vertical structure in the red giant branch of the \cmd.

@@

We select sources from our catalog within a $1~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \gaia\ within this window on the sky.
Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not at all obvious.
The posterior PDFs, shown in the bottom row, are more sharply peaked showing an increase in precision.
The posterior PDFs also have modes that are more similar to the previously determined distance to the cluster (vertical lines) for far more stars.
This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{distancesM67.png}
\caption{Proof of concept: parallax inference to the open cluster M67.
The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $1~{\rm deg}^2$ window centered on the position of M67.
Our prior, visualized as a mixture of Gaussians, is shown in green.
The upper right panel shows the \gaia\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
The posterior PDFs are much more clustered and sharply peaked around the ``true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
\label{fig:m67}
\end{figure}
